---
title: "Project Plan and EDA"
format:
    html:
        embed-resources: true
editor: visual
---

**Reading CSV**

```{r readingCSV}
fulldata <- read.csv("alzheimers_disease_data.csv")
fulldata$Diagnosis <- as.factor(fulldata$Diagnosis)
str(fulldata)
```

**Dropping columns**

```{r dropcolumns}
fulldata <- subset(fulldata, select = -c(PatientID, DoctorInCharge))
```

**Tree Model**

```{r tree}
library(tree)
treemodel <- tree(Diagnosis ~ ., data = fulldata)
plot(treemodel)
text(treemodel)
```

## Overview of the Problem

The data i

## Describe how this is a classification problem

## Provide some context about why this problem is interesting

## Dataset description

Describe the data (how many samples are there, are the columns of the data numeric, categorical) What challenges do you envision for your dataset (lots of missing values, high-dimensional data, etc.)

## Are there features that are highly correlated

## Evaluation metrics that are planned to be used

## Describe how you plan to evaluate your classification model
